---
title: "Actigraph RAW Data Screening"
author: "Tino Schneidewind"
date: "2025-06-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The aim of this document
The aim of this document is to investigate the actigraph data. I want to compare both the RAW output of all variables (1min resolution) to the cleaned (PVL, weartime validation) and aggregated (hourly) output. Through this I hope to identify if further cleaning is necessary, for example because the device is malfunctioning.

```{r load redcap, echo = FALSE, message=FALSE, warning=FALSE}
# libraries
library(readr);library(tidyverse);library(lubridate)

# week indicator
week_indicator = "week_1"
# load REDCap
redcap = read_csv("/Volumes/FS/_ISPM/CCH/Actual_Project/data/App_Personal_Data_Screening/redcap_data.csv") |>
  mutate(starttime = ymd_hms(starttime),
         endtime   = ymd_hms(endtime),
         redcap_event_name = substr(redcap_event_name, 13,18)) |>
  filter(redcap_event_name == week_indicator)|>
  filter(!(uid %in% c("ACT029U", "ACT034X", "ACT045O")))|>
  filter(uid != "ACT")

# unique series of uids
uids <- unique(redcap$uid)
```

### 1.chose a participant

```{r step1}
# chose a participant
uid_chosen <- uids[4] # options 2; 3; 4
loc <- paste0("/Volumes/FS/_ISPM/CCH/Actual_Project/data-raw/Actigraph/participants/", week_indicator, "/",uid_chosen, "/")
```

### 2. load all their variables

```{r load data, message=FALSE, warning=FALSE}
# weartime
WT_file <- list.files(loc, pattern = "validation_hourly")
WT_hourly <- read_csv(paste0(loc,WT_file)) |>
  mutate(datehour = timestamp_hour) |>
  select(datehour, ACC, validity)

# cardiac rythm
CR_file <- list.files(loc, pattern = "CR")
CR <- read_csv(paste0(loc,CR_file))

# heart rate
HR_file <- list.files(loc, pattern = "HR_RAW")
HR <- read_csv(paste0(loc,HR_file))

# heart rate variability
HRV_file <- list.files(loc, pattern = "HRV_RAW")
HRV <- read_csv(paste0(loc,HRV_file))

# inter beat interval
IBI_file <- list.files(loc, pattern = "IBI_RAW")
IBI <- read_csv(paste0(loc,IBI_file))

# steps
STEPS_file <- list.files(loc, pattern = "Steps")
STEPS <- read_csv(paste0(loc,STEPS_file))
```
### 3. aggregate to hourly, clean all variables by weartime validation and PVL

```{r step 3}
# cardiac rythm
CR_hourly <- CR |>
  mutate(datehour = floor_date(start_timestamp, unit = "hour")) |>
  group_by(datehour) |>
  summarize(bout_length_seconds_mean = mean(bout_length_seconds, na.rm = TRUE))

# heart rate
HR_hourly <- HR |>
  mutate(datehour = floor_date(timestamp, unit = "hour")) |>
  group_by(datehour) |>
  summarize(heart_rate_mean = mean(HeartRate, na.rm = TRUE))
  

# heart rate variability
HRV_hourly <- HRV |>
  mutate(datehour = floor_date(timestamp, unit = "hour")) |>
  group_by(datehour) |>
  summarize(breathing_rate_mean = mean(breathing_rate, na.rm = TRUE))
  
  
# inter beat interval
IBI_hourly <- IBI |>
  mutate(datehour = floor_date(timestamp, unit = "hour")) |>
  group_by(datehour) |>
  summarize(interbeat_interval_mean = mean(interbeat_interval, na.rm = TRUE))

# steps
STEPS_hourly <- STEPS |>
  mutate(datehour = floor_date(timestamp, unit = "hour")) |>
  group_by(datehour) |>
  summarize(steps_mean = mean(steps, na.rm = TRUE))
  


# create 1 dataframe for all variables from last to first PVL
actigraph <- data.frame(datehour = seq( 
  from = floor_date(redcap$starttime[redcap$uid == uid_chosen], unit = "hour"),
  to = floor_date(redcap$endtime[redcap$uid == uid_chosen], unit = "hour"),
  by = "hour"))

actigraph <- actigraph |>
  left_join(STEPS_hourly, by = "datehour") |>
  left_join(CR_hourly, by = "datehour") |>
  left_join(HR_hourly, by = "datehour") |>
  left_join(HRV_hourly, by = "datehour") |>
  left_join(IBI_hourly, by = "datehour") |>
  left_join(WT_hourly, by = "datehour") 

# set weartime validation filter
actigraph <- actigraph |>
  mutate(across(
    c(steps_mean, 
      bout_length_seconds_mean, 
      heart_rate_mean, 
      breathing_rate_mean, 
      interbeat_interval_mean, 
      ACC),
    ~ if_else(validity > 0, NA_real_, .)
  ))
```


### 4. visualize the cleaned data and the raw data

```{r plots, fig.align='center', fig.width=9, fig.height=15, echo=FALSE}
par(mfrow = c(6,2))

# CR
plot(CR$start_timestamp, CR$bout_length_seconds, pch = 19, col = "black", lwd = 0.5, main = "CR: bout length seconds")
plot(actigraph$datehour, actigraph$bout_length_seconds_mean, col = "red", main = "cleaned and aggregated")

# HR
plot(HR$timestamp, HR$HeartRate, pch = 19, col = "black", lwd = 0.5, main = "HR: heart rate")
plot(actigraph$datehour, actigraph$heart_rate_mean, col = "red")

# HRV
plot(HRV$timestamp, HRV$breathing_rate, pch = 19, col = "black", lwd = 0.5, main = "HRV: breathing rate")
plot(actigraph$datehour, actigraph$heart_rate_mean, col = "red")

# IBI
plot(IBI$timestamp, IBI$interbeat_interval, pch = 19, col = "black", lwd = 0.5, main = "IBI: interbeat interval")
plot(actigraph$datehour, actigraph$interbeat_interval_mean, col = "red")

# Steps
plot(STEPS$timestamp, STEPS$steps, pch = 19, col = "black", lwd = 0.5, main = "Steps")
plot(actigraph$datehour, actigraph$steps_mean, col = "red")

# ACC + validity
plot(WT_hourly$datehour, WT_hourly$ACC, col = "red", main = "ACC")
plot(WT_hourly$datehour, WT_hourly$validity, col = "red", main = "validity")
```

