---
title: "Week 1 Summary"
author: "Tino Schneidewind"
date: "2025-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prefix,echo=FALSE, warning=FALSE, message=FALSE}
rm(list=ls())


library(readr);library(tidyr);library(dplyr);library(readxl);library(knitr)
library(lubridate);library(stringr);library(ggplot2);library(gridExtra); library(grid)



```

## Week 1 overview of the personal environmental data
This report summarizes the variables humidity and temperature measure in the individuals house (IBH_HUM, IBH_TEMP), the humidity and temperature detected by worn necklaces (IBW_HUM, IBW_TEMP), the temperature of the device taped to the skin (IBT_TEMP) and the noise inside the individuals house (NS).

```{r data raw, echo = FALSE, warning=FALSE, message=FALSE}
# REDCap for uids and cutting the data
redcap = read_csv("../../data/redcap_data.csv") |>
  dplyr::mutate(starttime = ymd_hms(starttime),
                endtime   = ymd_hms(endtime),
                redcap_event_name = substr(redcap_event_name, 13,18))

# hourly summarized data
hourly_data = read_csv("../../data/week1_hourly_data.csv") |>
  mutate(hourly = format(hour, "%H"))

diurnal_cycle <- hourly_data |>
  group_by(hourly) |>
  summarise(IBH_HUM = mean(IBH_HUM, na.rm = TRUE),
            IBH_TEMP = mean(IBH_TEMP, na.rm = TRUE),
            IBW_HUM = mean(IBW_HUM, na.rm = TRUE),
            IBW_TEMP = mean(IBW_TEMP, na.rm = TRUE),
            IBT_TEMP = mean(IBT_TEMP, na.rm = TRUE),
            NS = mean(NS, na.rm = TRUE)) |>
  mutate(hourly = as.numeric(hourly)) |>
  na.omit()


# Calculate summary statistics (mean, median, sd) for all numeric columns
summary_manual <- data.frame(
  Variable = c("IBH_HUM", "IBH_TEMP", "IBW_HUM", "IBW_TEMP", "IBT_TEMP", "NS"),  
  mean = c(mean(hourly_data$IBH_HUM, na.rm = TRUE), mean(hourly_data$IBH_TEMP, na.rm = TRUE),
           mean(hourly_data$IBW_HUM, na.rm = TRUE), mean(hourly_data$IBW_TEMP, na.rm = TRUE),
           mean(hourly_data$IBT_TEMP, na.rm = TRUE), mean(hourly_data$NS, na.rm = TRUE)),
  median = c(median(hourly_data$IBH_HUM, na.rm = TRUE), median(hourly_data$IBH_TEMP, na.rm = TRUE),
             median(hourly_data$IBW_HUM, na.rm = TRUE), median(hourly_data$IBW_TEMP, na.rm = TRUE),
             median(hourly_data$IBT_TEMP, na.rm = TRUE), median(hourly_data$NS, na.rm = TRUE)),
  sd = c(sd(hourly_data$IBH_HUM, na.rm = TRUE), sd(hourly_data$IBH_TEMP, na.rm = TRUE),
         sd(hourly_data$IBW_HUM, na.rm = TRUE), sd(hourly_data$IBW_TEMP, na.rm = TRUE),
         sd(hourly_data$IBT_TEMP, na.rm = TRUE), sd(hourly_data$NS, na.rm = TRUE)))
```

### Distributions

```{r data distribtuione, echo = FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.height=4, fig.width=8}
plot1 <- ggplot(hourly_data, aes(x = IBH_HUM)) +
  geom_histogram(fill = "skyblue1", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "humidity [%]", y = "density", title = "House - Humidity") +
  lims(x = c(0,100)) +
  theme_classic()


plot2 <- ggplot(hourly_data, aes(x = IBW_HUM)) +
  geom_histogram(fill = "skyblue1", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "humidity [%]", y = "density", title = "Worn - Humidity") +
  lims(x = c(0,100)) +
  theme_classic()


plot3 <- ggplot(hourly_data, aes(x = IBH_TEMP)) +
  geom_histogram(fill = "brown3", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "temperature [°C]", y = "density", title = "House - Temperature") +
  lims(x = c(15,45)) +
  theme_classic()

plot4 <- ggplot(hourly_data, aes(x = IBW_TEMP)) +
  geom_histogram(fill = "brown3", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "temperature [°C]", y = "density", title = "Worn - Temperature") +
  lims(x = c(15,45)) +
  theme_classic()

plot5 <- ggplot(hourly_data, aes(x = NS)) +
  geom_histogram(fill = "grey", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "noise [dB]", y = "density", title = "House - Noise") +
  lims(x = c(15,90)) +
  theme_classic()

plot6 <- ggplot(hourly_data, aes(x = IBT_TEMP)) +
  geom_histogram(fill = "brown3", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "temperature [°C]", y = "density", title = "Taped - Temperature") +
  lims(x = c(15,45)) +
  theme_classic()

grid.arrange(plot1, plot2, plot5, plot3, plot4,  plot6, nrow = 2)

```

### Diurnal Cycle

```{r data diurnal, echo = FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.height=5.5, fig.width=7}
plot1 <- ggplot(diurnal_cycle, aes(x = hourly)) + 
  geom_line(aes(y = IBH_HUM, color = "House"), size = 1.2) + 
  geom_line(aes(y = IBW_HUM, color = "Worn"), size = 1.2) + 
  labs(
    x = "hour of the day", 
    y = "humidity [%]", 
    title = "Humidity",
    color = "Legend"  # Title for the legend
  ) +
  lims(y = c(20,50)) +
  scale_color_manual(values = c("House" = "skyblue", "Worn" = "skyblue4"),
                     breaks = c("Worn", "House")) + 
  theme_classic()


plot2 <- ggplot(diurnal_cycle, aes(x = hourly)) + 
  geom_line(aes(y = IBH_TEMP, color = "House"), size = 1.2) + 
  geom_line(aes(y = IBW_TEMP, color = "Worn"), size = 1.2) + 
  geom_line(aes(y = IBT_TEMP, color = "Taped"), size = 1.2) + 
  labs(
    x = "hour of the day", 
    y = "temperature [°C]", 
    title = "Temperature",
    color = "Legend"  # Title for the legend
  ) +
  lims(y = c(22,40)) +
  scale_color_manual(
    values = c("House" = "red", "Worn" = "brown3", "Taped" = "brown4"), 
    breaks = c("Taped", "Worn", "House")  # Set the desired legend order
  ) +
  theme_classic()


plot3 <- ggplot(diurnal_cycle, aes(x = hourly)) + 
  geom_line(aes(y = NS, color = "Noise"), size = 1.2) + 
  labs(
    x = "hour of the day", 
    y = "noise [dB]", 
    title = "Noise",
    color = "Legend"
  ) +
  lims(y = c(40,60)) +
    scale_color_manual(
    values = c("Noise" = "darkgrey")) +
  theme_classic()

grid.arrange(plot1, plot2, plot3, ncol = 1)
```

### Descriptive Statistics

```{r table stats, echo = FALSE}

kable(summary_manual, 
      format = "markdown",  
      col.names = c("Variable", "Mean", "Median", "Standard Deviation"))
```

### Scatterplots

```{r intra device, echo = FALSE, warning=FALSE, messages = FALSE, fig.align='center', fig.width=11, fig.height=4}

plot1 <- ggplot(hourly_data, aes(x = IBH_TEMP, y = IBH_HUM)) +
  geom_point(size = 0.7, color = "purple1") +
  labs(x = "temperature [°C]", y = "humidity [%]", title = "House") +
  theme_classic()

plot2 <- ggplot(hourly_data, aes(x = IBW_TEMP, y = IBW_HUM)) +
  geom_point(size = 0.7, color = "purple1") +
  labs(x = "temperature [°C]", y = "humidity [%]", title = "Worn") +
  theme_classic()

plot3 <- ggplot(hourly_data, aes(x = IBH_HUM, y = IBW_HUM)) +
  geom_point(size = 0.7, color = "skyblue2") +
  labs(x = "humidity [%]", y = "humidity [%]", title = "House x Worn") +
  theme_classic()

grid.arrange(plot1, plot2, plot3, nrow = 1)
```

<br>

```{r scatterl interdevice, echo=FALSE, fig.align='center', fig.height=6, fig.width=7, warning=FALSE, message=FALSE}
plot4 <- ggplot(hourly_data, aes(x = IBH_TEMP, y = IBT_TEMP)) +
  geom_point(size = 0.7, color = "brown2") +
  labs(x = "temperature [°C]", y = "temperature [°C]", title = "House x Taped") +
  theme_classic()


plot5 <- ggplot(hourly_data, aes(x = IBW_TEMP, y = IBT_TEMP)) +
  geom_point(size = 0.7, color = "brown2") +
  labs(x = "temperature [°C]", y = "temperature [°C]", title = "Worn x Taped") +
  theme_classic()



plot6 <- ggplot(hourly_data, aes(x = IBW_TEMP, y = IBH_TEMP)) +
  geom_point(size = 0.7, color = "brown2", alpha = 0.7) +
  labs(x = "temperature [°C]", y = "temperature [°C]", title = "Worn x House") +
  theme_classic()

plot7 <- ggplot(hourly_data, aes(x = IBH_TEMP, y = NS)) +
  geom_point(size = 0.7, color = "#884848") +
  labs(x = "temperature [°C]", y = "noise [dB]", title = "House") +
  theme_classic()


grid.arrange(plot4, plot5,  plot6, plot7, nrow = 2)
```

<br>